\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-01 paper/main.tex
title update
update values verification tables, update title

\title{\textbf{Advection versus diffusion in brain intraventricular transport}}

@@

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx with petsc4py as linear algebra backend
\cite{dolfinx2023preprint, Dalcin2011ParallelPython}.
The simulation code is openly
available\fixme{~\cite{zenodo-link-to-all-code}}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver
with block Jacobi preconditioning for the outer solves
and ILU preconditioning for the inner solves
\cite{fgmres, Jacobi1845UeberGleichungen, chan1997approximate}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \small
    \centering
    \caption{Errors of the velocity approximation $\uu_h$ in the
    $\mathbf{L}^2$ and $\mathbf{H}^1$ norms,
    the pressure $p_h$ in the $L^2$ norm,
    and the error in the maximum velocity magnitude $\uu_{h, {\mathrm{mag}}}$
    for mesh refinement on a cylinder mesh with the cilia-driven/no-cardiac flow model,
    with $\mu=1$.
    Additionally, we report the $L^2$ norm of the divergence of the velocity.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\normlinf{u_h} -\normlinf{u_{\mathrm{ex}}}$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.4e-01}  (--) 
        & \num{2.4e-02} (---) & \num{1.4e-07} \\

        1/2  & 41190   & \num{9.3e-03} (1.30) & \num{2.8e-01}  (0.61) & \num{3.7e-01} (0.53)
        & \num{5.0e-03} (2.25) & \num{2.8e-08} \\

        1/4  & 255016  & \num{3.5e-03} (1.42) & \num{1.4e-01}  (1.00) & \num{2.1e-01} (0.79)
        & \num{1.4e-03} (1.83) & \num{7.3e-11} \\

        1/8  & 1855337 & \num{1.3e-03} (1.46) & \num{6.9e-02}  (1.03) & \num{1.1e-01} (0.94)
        & \num{6.7e-04} (1.08) & \num{1.4e-10} \\   
        \bottomrule
    \end{tabular}
\end{table}

To verify the model implementation with the zebrafish brain ventricles mesh,
we compared velocity norms for varying degrees of refinement of the brain ventricles
mesh~(\Cref{tab:ventricles_norms_BDM}).
We used the original mesh obtained by meshing the surface geometry using
fTetWild~\cite{Hu2020FastWild}; the region around the photoconversion region is not
refined on the mesh used for the convergence study.
For the cilia traction vector $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ introduced in~\Cref{subsec:appendixA1},
we set a constant $\tau = 6.5\times 10^{-4}$ Pa and
$\lambda(\xx)=1$ in the numerical experiments.
We set $\rr = (1, 0, 1)$ in the ventral middle ventricle cilia region
and $\rr = -(1, 0, 1)$ in the dorsal middle ventricle and the anterior ventricle cilia
regions. We used $\gamma=10$ for the BDM penalty parameter.

For all meshes, we report the maximum discrete velocity magnitude $\normlinf{u_h}$.
We also report $\normlinf{u_h}$ scaled by the area $|\Gc|=\intGc{}$ of the cilia
boundary $\Gc$ over which tangential traction
$\btau$ was applied. We chose this measure using $|\Gc|$ as a scale because the regions
of the mesh marked as cilia-populated change slightly with every mesh refinement,
owing to the unstructured organization of the mesh cells.
These changes in the surface area over which the tangential traction is applied
result in changes to the maximum velocity magnitude,
since the maximum velocity magnitude is a result of the total amount of force
$\intGc{\sqrt{\btau\cdot \btau}}$ applied on the cilia boundary.

To account for the fact that the volume of the domain $|\Omega|=\intO{}$ also changes
slightly with mesh refinement, we report the $\mathbf{L}^2$ norm of $\uu_h$ scaled by $|\Omega|$.
Finally, we provide verification of mass conservation by reporting the $L^2$ norms of
the divergence $\nabla\cdot\uu_h$. We remark that the units used in the model
implementation was millimeters, which is why~\Cref{tab:ventricles_norms_BDM}
reports values in units of millimeters, instead of micrometers which has been used elsewhere in the paper. 
\begin{table}[!htbp]
    \centering
    \caption{Velocity norms under mesh refinement
    of the brain ventricles mesh using the cilia-driven/no-cardiac flow model
    with $\btau = \tau \lambda(\xx)P_{\nn}(\rr)$, using
    $\lambda(\xx)=1$,
    $\tau = 6.5\times 10^{-4}$ Pa,
    $\rr = (1, 0, 1)$ in the ventral middle ventricle and,
    $\rr = -(1, 0, 1)$ in the dorsal middle ventricle and
    the anterior ventricle.}\label{tab:ventricles_norms_BDM}
    \begin{tabular}{cc|cccc}
        \toprule
        \# cells & \# dofs &
        $\normlinf{u_h}$ $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\frac{1}{|\Gc|}\normlinf{u_h}$ $\left[\frac{1}{\mathrm{mm\cdot s}}\right]$
        & $\frac{1}{\vert\Omega\vert}\normltwovec{\uu_h}$
        $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\normltwo{\nabla\cdot\uu_h}$ $\left[\frac{\mathrm{mm^3}}{\mathrm{s^2}}\right]$\\
        \midrule 
        132134  & 943820  & 0.0276 & 0.97 & \num{3.6e-03} & \num{6.3e-16}\\
        180901  & 1289881 & 0.0284 & 0.98 & \num{3.7e-03} & \num{1.5e-13}\\
        262598  & 1870703 & 0.0292 & 0.98 & \num{3.8e-03} & \num{6.2e-16}\\
        395339  & 2811668 & 0.0292 & 0.96 & \num{3.8e-03} & \num{1.9e-14}\\
        627439  & 4448761 & 0.0286 & 0.92 & \num{3.7e-03} & \num{4.6e-14}\\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

\title{\Large\textbf{Advection versus diffusion in brain ventricular transport}}

@@

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx with petsc4py as linear algebra backend
\cite{dolfinx2023preprint, Dalcin2011ParallelPython}.
The simulation code is openly
available\fixme{~\cite{zenodo-link-to-all-code}}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver
with block Jacobi preconditioning \cite{fgmres, Jacobi1845UeberGleichungen}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \footnotesize
    \centering
    \caption{Error norms for the convergence study
    performed with a cylinder mesh with
    the cilia-driven/no-cardiac flow model using $\mu=1$.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses. The coarsest mesh had a mesh size $h_0$,
    and for every mesh refinement the mesh size $h$ was halved.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\vert\normlinf{u_h} - \normlinf{u_{\mathrm{ex}}}\vert$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.6e-01}  (--)  & \num{2.4e-02}  (--)  & \num{1.4e-07} \\  
        1/2  & 41190   & \num{9.3e-03} (1.30) & \num{2.8e-01}  (0.61) & \num{3.9e-01} (0.53) & \num{5.0e-03} (2.25) & \num{2.8e-08} \\  
        1/4  & 255016  & \num{3.5e-03} (1.42) & \num{1.4e-01}  (1.00) & \num{2.1e-01} (0.85) & \num{1.4e-03} (1.83) & \num{7.3e-11} \\     
        1/8  & 1855337 & \num{1.3e-03} (1.46) & \num{6.9e-02}  (1.03) & \num{1.1e-01} (0.92) & \num{6.7e-04} (1.08) & \num{1.5e-10} \\   
        \bottomrule
    \end{tabular}
\end{table}

To verify the model implementation with the zebrafish brain ventricles mesh,
we compared velocity norms for varying degrees of refinement of the brain ventricles
mesh~(\Cref{tab:ventricles_norms_BDM}).
We used the original mesh obtained by meshing the surface geometry using
fTetWild~\cite{Hu2020FastWild}; the region around the photoconversion region is not
refined on the mesh used for the convergence study.
In the numerical experiments, we used the cilia
traction vector $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$
defined in~\Cref{tab:r_vector_and_lambda_function}
and $\gamma=10$ for the BDM penalty parameter.

For all meshes, we report the maximum discrete velocity magnitude $\normlinf{u_h}$.
We also report $\normlinf{u_h}$ scaled by the total force $F=\intGc{\sqrt{\btau\cdot \btau}}$
applied on the cilia boundary $\Gc$. We chose this measure using $F$ as a scale because the regions
of the mesh marked as cilia-populated change slightly with every mesh refinement,
owing to the unstructured organization of the mesh cells.
These changes in the surface area over which the tangential traction is applied
result in changes to $\normlinf{u_h}$,
since the maximum velocity magnitude is a result of the total amount of force
$F$ applied on the cilia boundary.

To account for the fact that the volume of the domain $|\Omega|=\intO{}$ also changes
slightly with mesh refinement, we report the $\mathbf{L}^2$ norm of $\uu_h$ scaled by $|\Omega|$.
Finally, we provide verification of mass conservation by reporting the $L^2$ norms of
the divergence $\nabla\cdot\uu_h$. We remark that the units used in the model
implementation was millimeters, which is why~\Cref{tab:ventricles_norms_BDM}
reports values in units of millimeters, instead of micrometers which has been used elsewhere in the paper. 
\begin{table}[!htbp]
    \centering
    \small
    \caption{Velocity norms under mesh refinement
    of the brain ventricles mesh using the cilia-driven/no-cardiac flow model.}
    \label{tab:ventricles_norms_BDM}
    \begin{tabular}{cc|cccc}
        \toprule
        \# cells & \# dofs &
        $\normlinf{u_h}$ $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\frac{1}{F}\normlinf{u_h}$ $\left[\frac{1}{\mathrm{Pa\cdot mm\cdot s}}\right]$
        & $\frac{1}{\vert\Omega\vert}\normltwovec{\uu_h}$
        $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\normltwo{\nabla\cdot\uu_h}$ $\left[\frac{\mathrm{mm^3}}{\mathrm{s^2}}\right]$\\
        \midrule 
        132134  & 943820  & 0.026 & \num{3.8e3} & \num{2.2e-03} & \num{1.9e-16}\\
        180901  & 1289881 & 0.027 & \num{3.7e3} & \num{2.2e-03} & \num{3.0e-16}\\
        262598  & 1870703 & 0.027 & \num{3.7e3} & \num{2.3e-03} & \num{6.7e-16}\\
        395339  & 2811668 & 0.028 & \num{3.6e3} & \num{2.2e-03} & \num{2.0e-14}\\
        627439  & 4448761 & 0.027 & \num{3.5e3} & \num{2.2e-03} & \num{4.9e-14}\\
        \bottomrule
    \end{tabular}
\end{table}
